---
title: "survivorship"
author: "allyson_demerlis"
date: "2023-04-17"
output: html_document
---

# Load packages and data
```{r}
library(tidyverse)
library(survival)
library(ggpubr)
library(survminer)
library(coxme)
library(ggsurvfit)
library(gtsummary)

survivorship <- readxl::read_xlsx("survivorship_data.xlsx", sheet = "kaplan meier data")
survivorship$Genet <- as.factor(survivorship$Genet)
```
# Definitions 

Kaplan-Meier survivorship curves are built off of data that use binary encoding for time-to-event and censoring.
Time-to-event is the number of days until the event happened.
In my case, it would be time to mortality. 

1 = event (mortality)
0 = censored

Censoring occurs if a subject has not experienced the event of interest by the end of data collection. 

Reference: https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html 


# Proportion survived bleaching





# Kaplan-Meier Survivorship: All time points

## Kaplan-Meier survivorship curve grouped by transplant
```{r}
Surv(survivorship$DaysToDeath, survivorship$StatusAtEnd)

s1 <- survfit(Surv(DaysToDeath, StatusAtEnd) ~ Transplant, data = survivorship)
str(s1)

survfit2(Surv(DaysToDeath, StatusAtEnd) ~ Transplant, data = survivorship) %>% 
  ggsurvfit() +
  labs(x = "Days", y = "Overall survival probability") +
  scale_ggsurvfit() +
  add_pvalue(caption = "Log-rank {p.value}") +
  add_pvalue(location  = "annotation", x = 15)
```

## adjusted Cox Proportion Hazards Regression model to include Genet 
```{r}
coxph(Surv(DaysToDeath, StatusAtEnd) ~ Genet + strata(Transplant), data = survivorship) %>%
  survfit2() %>%
  ggsurvfit()

#~ Genet + strata(Transplant): This part of the formula specifies the independent variables.
#Genet is a variable in the dataset. It could be a continuous variable, a categorical variable, 
#or any other type of variable you're investigating to see how it affects the survival time.
# strata(Transplant) is used to indicate that the model should stratify by the Transplant variable. Stratification here means the analysis is done separately within each stratum of Transplant, allowing for different baseline hazards in different strata but keeping the effect of Genet constant across strata.

# Cox proportional hazards model to analyze how the variable Genet affects the survival time (DaysToDeath), while controlling for different levels of Transplant through stratification. 
```
## Assessing proportional hazards

### Checking assumptions are met 

One assumption of the Cox proportional hazards regression model is that the hazards are proportional at each point in time throughout follow-up. The cox.zph() function from the {survival} package allows us to check this assumption. 
```{r}
mv_fit <- coxph(Surv(DaysToDeath, StatusAtEnd) ~ Genet + Transplant, data = survivorship)
cz <- cox.zph(mv_fit)
print(cz) # Genet and transplant are not significant, so they meet the assumptions
```

### Estimating x-year survival

One quantity often of interest in a survival analysis is the probability of surviving beyond a certain number of years, x.
```{r}
summary(survfit(Surv(DaysToDeath, StatusAtEnd) ~ Transplant, data = survivorship), times = 350) #350 days in 
```

### Estimating median survival time

Another quantity often of interest in a survival analysis is the average survival time, which we quantify using the median. Survival times are not expected to be normally distributed so the mean is not an appropriate summary.
```{r}
survfit(Surv(DaysToDeath, StatusAtEnd) ~ Transplant, data = survivorship)
```



### Comparing survival times between groups

We can conduct between-group significance tests using a log-rank test. The log-rank test equally weights observations over the entire follow-up time and is the most common way to compare survival times between groups. There are versions that more heavily weight the early or late follow-up that could be more appropriate depending on the research question (see ?survdiff for different test options).

We get the log-rank p-value using the survdiff function.
```{r}
survdiff(Surv(DaysToDeath, StatusAtEnd) ~ Transplant, data = survivorship)
```

### cox hazards ratio statistics
```{r}
coxph(Surv(DaysToDeath, StatusAtEnd) ~ Transplant, data = survivorship) %>% 
  tbl_regression(exp = TRUE) 
```


Following Maddie Kaufman's code:
```{r}
# km survivorship curves ----
# Kaplan-Meier is univariate; Cox proportional hazards is multivariable
#null model: no random effects or treatment
model0=coxph(Surv(DaysToDeath, StatusAtEnd) ~ 1, data = survivorship)
summary(model0)

#model without random effects but including treatment
model1=coxph(Surv(DaysToDeath, StatusAtEnd) ~ Transplant, data = survivorship) 
summary(model1)

model2 = coxme(Surv(DaysToDeath, StatusAtEnd) ~ Transplant + (1|Genet), data = survivorship)
summary(model2)

# create survival object using the Kaplan-Meier method
surv <- Surv(time = survivorship$DaysToDeath, event = survivorship$StatusAtEnd)
surv

# run survival model for each species
fit_reef <- survfit(surv ~ Transplant, data = survivorship)
summary(fit_reef)

ggsurvplot(fit_reef, data = survivorship, risk.table = TRUE, pval = TRUE, conf.int=TRUE)
```

LRT function
```{r}
## Model comparison function

#two *nested* models as input, in this case obj1 is the model without the predictor variable and obj2 is the model with the predictor variable 
lrt <- function (obj1, obj2) {
  L0 <- logLik(obj1)
  L1 <- logLik(obj2)
  L01 <- as.vector(- 2 * (L0 - L1))
  df <- attr(L1, "df") - attr(L0, "df")
  list(L01 = L01, df = df,
       "p-value" = pchisq(L01, df, lower.tail = FALSE))
}
```


Compare three models to see which one is best fit of data.
```{r}
#Compare mixed effects model versus only fixed effects model. If not significant, go with simpler model.
anova(model0, model1, model2)
```



Following Michael Studivan's code: (https://github.com/mstudiva/PortMiami-urban-coral-resilience/blob/main/coralrelocation/fisher%20survivorship.R)
```{r}
#### packages ####

library(ggplot2)
library(ggpubr)
library(rcompanion)
library(MASS)
library(stringr)
library(survival)
library(survminer)
library(dplyr)

#### survivorship by species ####

# create survival object using the Kaplan-Meier method
surv <- Surv(time = survivorship$DaysToDeath, event = survivorship$StatusAtEnd)
surv

# run survival model for each transplantation direction
fit_reef <- survfit(surv ~ Transplant, data = survivorship)
summary(fit_reef)

# Kaplan-Meier plots for each species
fill.color<-c("#43a2ca", "#a97c50","#addd8e", "#006837")

survival_transplant<-ggsurvplot(fit_reef, data = survivorship, pval = TRUE, xlab="Days", ylab="Survival Probability",
                                    conf.int = T, risk.table=T, palette=fill.color,
                                    break.time.by=31, xlim=c(0,372), risk.table.y.text = FALSE) + ggtitle("Species") 
survival_transplant

# hazard ratio by species
hazard_transplant <- coxph(surv ~ Transplant, data = survivorship)
summary(hazard_transplant)

# hazard ratio plot
hazplot_transplant <- ggforest(hazard_transplant, data = survivorship)
hazplot_transplant


# run survival model for each transplantation direction with genet included
fit_reefgenet <- survfit(surv ~ Genet+Transplant, data = survivorship)
summary(fit_reefgenet)

# Kaplan-Meier plot
survival_transplantgenet<-ggsurvplot(fit_reefgenet, data = survivorship, pval = TRUE, xlab="Days", ylab="Survival Probability",
                                     xlim=c(0,372), risk.table.y.text = FALSE)
survival_transplantgenet

# hazard ratio by species
hazard_transplant <- coxph(surv ~ Transplant, data = survivorship)
summary(hazard_transplant)

# hazard ratio plot
hazplot_transplant <- ggforest(hazard_transplant, data = survivorship)
hazplot_transplant

#### multiplot #####

# survivorship by species and size figure
multiplot<-ggarrange(survival_transplant$plot,
                                     hazplot_transplant,
                                     heights = c(2, 0.5, 0.75),
                                     ncol = 2, nrow = 1)

ggsave("survivorship.pdf", multiplot, width=20, height=6,dpi = 300)
```



Coxme is the cox mixed-effects version of the cox proportional hazards model, and can include co-variates.
```{r}
model <- coxme(Surv(DaysToDeath, StatusAtEnd) ~ Transplant + (1|Genet), data = survivorship)
summary(model)

coxph(Surv(DaysToDeath, StatusAtEnd) ~ Genet + Transplant, data = survivorship)

summary(coxph(Surv(DaysToDeath, StatusAtEnd) ~ Genet + Transplant, data = survivorship))
```



# Kaplan-Meier Survivorship: Only E-M and M-M
```{r}
survivorship %>% 
  filter(Transplant == "Emerald -> MacN" | Transplant == "MacN -> MacN") -> surv_macN

Surv(surv_macN$DaysToDeath, surv_macN$StatusAtEnd)

s2 <- survfit(Surv(DaysToDeath, StatusAtEnd) ~ Transplant, data = surv_macN)
str(s2)

survfit2(Surv(DaysToDeath, StatusAtEnd) ~ Transplant, data = surv_macN) %>% 
  ggsurvfit() +
  labs(x = "Days", y = "Overall survival probability") +
  scale_ggsurvfit() +
  add_pvalue(caption = "Log-rank {p.value}") +
  add_pvalue(location  = "annotation", x = 15)

coxph(Surv(DaysToDeath, StatusAtEnd) ~ Genet + strata(Transplant), data = surv_macN) %>%
  survfit2() %>%
  ggsurvfit() 

summary(coxph(Surv(DaysToDeath, StatusAtEnd) ~ Genet + strata(Transplant), data = surv_macN))

coxme(Surv(DaysToDeath, StatusAtEnd) ~ Transplant + (1|Genet), data = surv_macN)
```

# Kaplan-Meier Survivorship: removing first time point
```{r}
survivorship %>%
  filter(DaysToDeath != 51) -> surv_AprNov

survfit2(Surv(DaysToDeath, StatusAtEnd) ~ Transplant, data = surv_AprNov) %>% 
  ggsurvfit() +
  labs(x = "Days", y = "Overall survival probability") +
  scale_ggsurvfit() +
  add_pvalue(caption = "Log-rank {p.value}") +
  add_pvalue(location  = "annotation", x = 15)
  
summary(survfit(Surv(DaysToDeath, StatusAtEnd) ~ Transplant, data = surv_AprNov), times = 350)

coxph(Surv(DaysToDeath, StatusAtEnd) ~ Genet + strata(Transplant), data = surv_AprNov) %>%
  survfit2() %>%
  ggsurvfit() 

summary(coxph(Surv(DaysToDeath, StatusAtEnd) ~ Genet + strata(Transplant), data = surv_AprNov))

coxme(Surv(DaysToDeath, StatusAtEnd) ~ Transplant + (1|Genet), data = surv_AprNov)
```

