---
title: "environmental_data_plots"
author: "allyson_demerlis"
date: "2023-08-31"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
library(cowplot)
library(ggpubr)
library(openair)
library(seacarb)
library(scales)   # to access breaks/formatting functions
library(gridExtra) # for arranging plots
library(ggradar)
library(doBy)
library(reshape2)
library(zoo)
```

Mike Jankulak created composite files of all time points for each site for each instrument. Some instruments were swapped between sites, so he made sure to group the sites correctly. There is a metadata file that shows all the instrument deployments.

# Temperature

## import STR files
```{r}
str_emerald <- read_excel("raw_data/str-emerald-sn5042.xlsx", sheet = "str-emerald-sn5042")

str_emerald$`Datetime (UTC)` <- as.POSIXct(str_emerald$`Datetime (UTC)`, tz="UTC")
str_emerald$`Datetime (EST)` <- with_tz(str_emerald$`Datetime (UTC)`, "America/New_York")

str_emerald %>% 
  mutate(reef = "Emerald") %>% 
  dplyr::select(!Time) %>% 
  dplyr::select(!`Datetime (UTC)`) %>% 
  dplyr::select(!`Sample Number`) -> str_emerald
  
str_macnorth <- read_excel("raw_data/str-macnorth-sn9235.xlsx", sheet = "str-macnorth-sn9235")

str_macnorth$Datetime <- as.POSIXct(str_macnorth$Datetime, tz="UTC")
str_macnorth$`Datetime (EST)` <- with_tz(str_macnorth$Datetime, "America/New_York")

str_macnorth %>% 
  mutate(reef = "MacNorth") %>% 
  dplyr::select(!Datetime) %>% 
  dplyr::select(!`Sample Number`) %>% 
  dplyr::select(!Time) -> str_macnorth
```


## just STR data
```{r}
full_join(str_emerald, str_macnorth) -> str_all

# Identify the lowest and highest points
lowest_point <- str_all %>% filter(Temperature == min(Temperature))
highest_point <- str_all %>% filter(Temperature == max(Temperature))

str_all %>% filter(reef == "Emerald") %>% filter(Temperature == max(Temperature))

str_all %>% 
  filter(Date >= "2022-12-06") %>% 
ggplot(., aes(x=`Datetime (EST)`, y=Temperature, color = reef)) +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "Temperature (ºC)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) +
  geom_hline(yintercept = 31, linetype = "dashed", color = "black") +
  annotate("text", x = as.POSIXct("2023-11-20"), y = 31, label = "31ºC", vjust = -1) +
  geom_point(data = lowest_point, aes(`Datetime (EST)`, Temperature), color = "#5370c9", size = 2) +
  geom_text(data = lowest_point, aes(`Datetime (EST)`, Temperature, label = "17.89ºC"), vjust = 1.1, hjust = -0.1, color = "black") +
  geom_point(data = highest_point, aes(`Datetime (EST)`, Temperature), color = "#5370c9", size = 2) +
  geom_text(data = highest_point, aes(`Datetime (EST)`, Temperature, label = "33.55ºC"), vjust = 0, hjust = -0.2, color = "black") +
labs(title = "2022-2023 Temperature Data at Emerald Reef and MacArthur Causeway", x="Month") 

  #annotate("text", x = as.POSIXct("2023-09-30"), y = 18, label= "STR data at Emerald stopped 9/9", colour = "black", hjust = .56)
```

## append CTD temperature data for Emerald after 9/9/23
```{r}
ctd_emerald <- read_excel("raw_data/ctd-emerald-composite.xlsx", sheet = "ctd-emerald-composite")

ctd_emerald %>% 
  dplyr::select(`Timestamp UTC`, Temperature) -> ctd_emerald_temp

ctd_emerald_temp$`Timestamp (UTC)` <- as.POSIXct(ctd_emerald_temp$`Timestamp UTC`, tz="UTC")
ctd_emerald_temp$`Datetime (EST)` <- with_tz(ctd_emerald_temp$`Timestamp (UTC)`, "America/New_York")

ctd_emerald_temp %>%
  mutate(Datetime_EST = `Datetime (EST)`) %>% 
  separate(Datetime_EST, into = c("Date", "Time"), sep = " ") %>% 
  mutate(Date = ymd(Date)) %>% 
  dplyr::select(-1,-3) %>% 
  mutate(reef = "Emerald") %>% 
  filter(Date >= "2023-09-09") -> ctd_emerald_temp_sept9

str_all %>% 
  filter(Date >= "2022-12-06") %>% 
  full_join(ctd_emerald_temp_sept9) %>% 
ggplot(., aes(x=`Datetime (EST)`, y=Temperature, color = reef)) +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "Temperature (ºC)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) +
  geom_hline(yintercept = 31, linetype = "dashed", color = "black") +
  annotate("text", x = as.POSIXct("2023-11-20"), y = 31, label = "31ºC", vjust = -1) +
  geom_point(data = lowest_point, aes(`Datetime (EST)`, Temperature), color = "#5370c9", size = 2) +
  geom_text(data = lowest_point, aes(`Datetime (EST)`, Temperature, label = "17.89ºC"), vjust = 1.1, hjust = -0.1, color = "black") +
  geom_point(data = highest_point, aes(`Datetime (EST)`, Temperature), color = "#5370c9", size = 2) +
  geom_text(data = highest_point, aes(`Datetime (EST)`, Temperature, label = "33.55ºC"), vjust = 0, hjust = -0.2, color = "black") +
labs( x="Month") +
  theme(text = element_text(size = 12)) 
#ggsave("temperatureplot.pdf", width = 6, height = 4)

str_all %>% 
  filter(Date >= "2022-12-06") %>% 
  full_join(ctd_emerald_temp_sept9) %>% 
ggplot(., aes(x=`Datetime (EST)`, y=Temperature, color = reef)) +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "Temperature (ºC)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) +
labs( x="Month") +
  theme(text = element_text(size = 12)) 
#ggsave("temperatureplot_withMMM.png")
```

## temp summary

### get MMM from CRW
```{r}
library(raster)
library(parallel)
library(ncdf4)


  # Read climatology file
  Climatology_MMM <- "ct5km_climatology_v3.1.nc"
  MMM.data <- brick(Climatology_MMM) 
  #MMM.data
  
  # macNorth
  lat_m <- (25.772927)
  lon_m <- (-80.15263)
  extract.pts_macn <- cbind(lon_m,lat_m)
  
  #emerald reef 
  lat_em <- (25.67735)
  lon_em <- (-80.097383)
  extract_em <- cbind(lon_em, lat_em)
  
  
  
  #MMM <- raster::extract(MMM.data, extract.pts,method="bilinear")
  MMM_mac <- raster::extract(MMM.data, extract.pts_macn,method="simple") #29.5
  MMM_emerald <- raster::extract(MMM.data, extract_em,method="simple") #29.5

  #write.csv(MMM_mac, "macnorth_CRW_MMM_1985-2012.csv", row.names=FALSE)

 # write.csv(MMM_emerald, "emerald_CRW_MMM_1985-2012.csv", row.names=FALSE)

```


add season 
```{r}
all_temp_data %>% 
  mutate(Season = case_when(`Datetime (EST)` < "2023-05-01 00:00:00" ~ "Dry",
                            `Datetime (EST)` >= "2023-05-01 00:00:00" ~ "Wet")) -> all_temp_data
```


### temp summary for Site 
```{r}
str_all %>% 
  filter(Date >= "2022-12-06") %>% 
  full_join(ctd_emerald_temp_sept9) %>% 
  dplyr::select(!Time) -> all_temp_data

temp.summary<- summaryBy(Temperature ~ reef, data = all_temp_data,
                         FUN = function(x) { c(max = max(x), min = min(x)) } )

temp.summary$Temp.seasonal <- temp.summary$Temperature.max - temp.summary$Temperature.min

# Daily Range
Daily.range <- summaryBy(Temperature ~ Date + reef, data=all_temp_data, FUN=c(max,min,mean,sd))

Daily.range$temp.daily.range <- Daily.range$Temperature.max - Daily.range$Temperature.min

temp.d.range <- summaryBy(temp.daily.range ~ reef, data = Daily.range,
                          FUN = function(x) { c(mean = mean(x)) } )
Daily.range %>% 
  filter(Temperature.max > 30.5) %>% 
  group_by(reef) %>% 
  dplyr::summarise(d.above.30.5 = n()) -> temp.d.30.5

full_join(temp.summary, temp.d.range) %>% 
  full_join(., temp.d.30.5) -> temp_summary
```

### DHW
```{r}
all_temp_data %>% 
  dplyr::group_by(Date, reef) %>% 
  mutate(mean_daily_temp = mean(Temperature)) %>% 
  ungroup() %>% 
  dplyr::select(!c("Temperature","Datetime (EST)")) %>% 
  distinct() %>% 
  mutate(HotSpot = case_when(reef == "MacNorth" ~ (mean_daily_temp - 29.55),
                             reef == "Emerald" ~ (mean_daily_temp - 29.53))) %>% 
  mutate(HotSpot = ifelse(HotSpot >= 0, HotSpot, 0)) %>% 
  mutate(Stress = ifelse(HotSpot >= 1, HotSpot, 0)) %>% 
  mutate(Weeks_Stress = Stress/7)  -> Long.data.DHW

Long.data.DHW %>% 
  filter(reef == "Emerald") -> DHW_emerald

DHW_emerald$DHW<-rollapplyr(DHW_emerald$Weeks_Stress,list(-(83:0)),sum,fill=NA)

Long.data.DHW %>% 
  filter(reef == "MacNorth") -> DHW_mac

DHW_mac$DHW<-rollapplyr(DHW_mac$Weeks_Stress,list(-(83:0)),sum,fill=NA)

rbind(DHW_emerald, DHW_mac) %>% 
ggplot(., aes(x=Date)) + 
    geom_hline(yintercept=8, linetype="dashed")+
    geom_hline(yintercept=4, linetype="dashed")+
    geom_line(aes(y = DHW, color = reef)) +
    ylab("DWH (C-week)  /  Temperature (C)") +
    xlab("Date") +
    theme_classic()  +
  scale_color_manual(values = c("#edb9c3", "#5370c9"))
#ggsave("DHW_linegraph.pdf")

# Highest DWH per year
rbind(DHW_emerald, DHW_mac) %>% 
          group_by(reef) %>%
          summarize(max = max(DHW, na.rm = TRUE)) %>% 
          rename(maxDHW=max) -> MaxDHW

```

```{r}
cbind(temp_summary, MaxDHW) %>% 
  dplyr::select(-7) -> temp_summary_DHW

write_csv(temp_summary_DHW, "temp_DHW_summary.csv")
```

## Radar plot
```{r}
scale_by_max <- function(x) {
  x * 100 / max(x, na.rm = TRUE)
}

 temp_summary_DHW %>%
   column_to_rownames(var="reef") %>% 
  dplyr::mutate(across(everything(), as.numeric)) %>%
  dplyr::mutate(across(everything(), scale_by_max)) -> temp_DHW_summary_percent

# To use the fmsb package, I have to add 2 lines to the dataframe: the max and min of each variable to show on the plot!
temp_DHW_summary_percent <- rbind(rep(100,5), rep(0,5), temp_DHW_summary_percent)

# Color vector
colors_border=c("#edb9c3", "#5370c9")
colors_in=c("#edb9c3", "#5370c9")

# Define colors with transparency using adjustcolor
colors_border <- c(adjustcolor("#edb9c3", alpha = 0.9), adjustcolor("#5370c9", alpha = 0.9))
colors_in <- c(adjustcolor("#edb9c3", alpha = 0.4), adjustcolor("#5370c9", alpha = 0.4))

library(fmsb)
# plot with default options:
#pdf("radarchart_withDHW.pdf")
radarchart(temp_DHW_summary_percent, axistype=1, 
            #custom polygon
            pcol=colors_border , pfcol=colors_in , plwd=4 , plty=1,
            #custom the grid
            cglcol="grey", cglty=1, axislabcol="grey", caxislabels=seq(0,100,25), cglwd=0.8,
            #custom labels
            vlcex=0.9
)
```



# PAR

```{r}
ecopar_macnorth = readxl::read_xlsx("raw_data/ecopar-macnorth-composite.xlsx", sheet = "ecopar-macnorth-composite")
ecopar_emerald = readxl::read_xlsx("raw_data/ecopar-emerald-composite.xlsx", sheet = "ecopar-emerald-composite")
#10 measurements were taken instantaneously at every 30 min interval for each day.
# PAR(avg) is the average over the 10 measurements at each interval

ecopar_macnorth$`Datetime (EST)` <- with_tz(ecopar_macnorth$`Datetime (UTC)`, "America/New_York")
ecopar_emerald$`Datetime (EST)` <- with_tz(ecopar_emerald$`Datetime (UTC)`, "America/New_York")

ecopar_macnorth %>% 
  dplyr::select(!c("PAR (calc)", "Date", "Time", "Datetime (UTC)")) %>% 
  drop_na() %>% 
  filter(`Datetime (EST)` >= as.Date("2022-12-06")) %>% 
  mutate(reef = "MacNorth") -> ecopar_macnorth_clean

ecopar_emerald %>% 
  dplyr::select(!c("PAR (calc)", "Date", "Time", "Datetime (UTC)")) %>% 
  drop_na() %>% 
  filter(`Datetime (EST)` >= as.Date("2022-12-06")) %>% 
  mutate(reef = "Emerald") -> ecopar_emerald_clean
  
rbind(ecopar_emerald_clean, ecopar_macnorth_clean) %>% 
  dplyr::select(`Datetime (EST)`, `PAR(avg)`, reef) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  group_by(Date, reef) %>% 
  summarise(daily_dose = sum(`PAR(avg)`)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) -> PAR_summary

 parplot<- ggplot(PAR_summary, aes(x= Date , y=PAR_mol_day, color = reef)) +
  theme_classic() +
  geom_line() + 
  labs(x="Month", y = "PAR (mol/day)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) + 
  theme(text = element_text(size = 15))+
   theme(legend.position = "none")
#i think the PAR meter failed in July - Oct for MacNorth
#ggsave("PAR_wholetimeset.pdf")
```

```{r}
rbind(ecopar_emerald_clean, ecopar_macnorth_clean) %>% 
  select(`Datetime (EST)`, `PAR(avg)`, reef) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  mutate(Season = case_when(Date < "2023-05-01" ~ "Dry",
                            Date >= "2023-05-01" ~ "Wet")) %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  group_by(Date, reef, Season) %>% 
  summarise(daily_dose = sum(`PAR(avg)`)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  ungroup() %>% 
  group_by(reef, Season) %>% 
  summarise(count=n(), mean_PARmolday=mean(PAR_mol_day), sd_PARmolday=sd(PAR_mol_day), min_PARmolday= min(PAR_mol_day), max_PARmolday=max(PAR_mol_day), range_PARmolday = max_PARmolday-min_PARmolday) %>% 
  write_csv("PAR_molperday_summary.csv")
```

```{r}
rbind(ecopar_emerald_clean, ecopar_macnorth_clean) %>% 
  select(`Datetime (EST)`, `PAR(avg)`, reef) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  mutate(Season = case_when(Date < "2023-05-01" ~ "Dry",
                            Date >= "2023-05-01" ~ "Wet")) %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  group_by(Date, reef, Season) %>% 
  summarise(daily_dose = sum(`PAR(avg)`)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>%  
  ggplot(., aes(x= Season , y=PAR_mol_day, fill = reef)) +
  theme_classic() +
  geom_boxplot() +
  labs(x="Season", y = "PAR (mol/day)") +
  scale_fill_manual(values = c("#edb9c3", "#5370c9")) + 
  theme(text = element_text(size = 15))
ggsave("PAR_boxplot_season.pdf")
```


# Salinity

```{r}
ctd_emerald <- read_xlsx("raw_data/ctd-emerald-composite.xlsx", sheet = "ctd-emerald-composite")

ctd_emerald$`Datetime (EST)` <- with_tz(ctd_emerald$`Timestamp UTC`, "America/New_York")

ctd_emerald %>% 
  dplyr::select(`Datetime (EST)`, Salinity) %>% 
  mutate(reef = "Emerald") %>% 
  filter(`Datetime (EST)` >= as.Date("2022-12-06")) -> ctd_emerald_salinity

ctd_macnorth <- read_xlsx("raw_data/ctd-macnorth-composite.xlsx", sheet = "ctd-macnorth-composite")

ctd_macnorth$`Datetime (EST)` <- with_tz(ctd_macnorth$`Timestamp UTC`, "America/New_York")

ctd_macnorth %>% 
  select(`Datetime (EST)`, Salinity) %>% 
  mutate(reef = "MacNorth") %>% 
  filter(`Datetime (EST)` >= as.Date("2022-12-06")) -> ctd_macnorth_salinity
```

```{r}
rbind(ctd_emerald_salinity, ctd_macnorth_salinity)  -> ctd_salinity
```


```{r}
rbind(ctd_emerald_salinity, ctd_macnorth_salinity) %>% 
  mutate(Season = case_when(`Datetime (EST)` < as.Date("2023-05-01") ~ "Dry",
                            `Datetime (EST)` >= as.Date("2023-05-01") ~ "Wet")) %>% 
  group_by(reef, Season) %>% 
  ggplot(., aes(x= `Datetime (EST)` , y=Salinity, color = reef)) +
  theme_classic() +
  geom_line() +
  labs(x="Date", y = "Salinity") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) + 
  theme(text = element_text(size = 15))
#ggsave("salinity_continuous.pdf")

rbind(ctd_emerald_salinity, ctd_macnorth_salinity) %>% 
  mutate(Season = case_when(`Datetime (EST)` < as.Date("2023-05-01") ~ "Dry",
                            `Datetime (EST)` >= as.Date("2023-05-01") ~ "Wet")) %>% 
  group_by(reef, Season) %>% 
  ggplot(., aes(x= Season , y=Salinity, fill = reef)) +
  theme_classic() +
  geom_boxplot() +
  labs(x="Season", y = "Salinity") +
  scale_fill_manual(values = c("#edb9c3", "#5370c9")) + 
  theme(text = element_text(size = 15)) 
ggsave("salinity_boxplot.pdf")
```

there are some definite outliers. remove those and re-plot
```{r}
# Define the window size (e.g., for a couple of days, window_size = 2)
window_size <- 3

# Calculate the rolling maximum temperature within the specified range
data <- ctd_salinity %>%
  separate(`Datetime (EST)`, into = c("Date", "Time"), sep = " ") %>% 
  dplyr::group_by(Date, reef) %>% 
  dplyr::mutate(mean_daily_salinity = mean(Salinity)) %>% 
  arrange(Date) %>%
  mutate(RollingMax = rollmax(mean_daily_salinity, window_size, fill = NA, align = "right"),
         RollingMin = rollapply(mean_daily_salinity, window_size, min, fill = NA, align = "right"),
         MaxThreshold = RollingMax + 1,
         MinThreshold = RollingMin - 1)

# Filter out temperatures exceeding the threshold (+3 degrees from the rolling max)
filtered_data <- data %>%
  filter(Salinity <= MaxThreshold & Salinity >= MinThreshold)

filtered_data %>% 
  ggplot(., aes(x=as.Date(Date), y=Salinity, color=reef)) +
  geom_point()
  #ok this looks better
```

```{r}
salinity_plot<- ggplot(filtered_data, aes(x= as.Date(Date), y=Salinity, color = reef)) +
  theme_classic() +
  geom_line() +
  labs(x="Date", y = "Salinity") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) + 
  theme(text = element_text(size = 15)) +
  theme(legend.position = "none")
#ggsave("salinity_continuous.pdf")
```



# Depth

```{r}
ctd_emerald <- read_xlsx("raw_data/ctd-emerald-composite.xlsx", sheet = "ctd-emerald-composite")

ctd_emerald$`Datetime (EST)` <- with_tz(ctd_emerald$`Timestamp UTC`, "America/New_York")

ctd_emerald %>% 
  select(`Datetime (EST)`, Depth) %>% 
  mutate(reef = "Emerald") %>% 
  filter(`Datetime (EST)` >= as.Date("2022-12-06")) -> ctd_emerald_depth

ctd_macnorth <- read_xlsx("raw_data/ctd-macnorth-composite.xlsx", sheet = "ctd-macnorth-composite")

ctd_macnorth$`Datetime (EST)` <- with_tz(ctd_macnorth$`Timestamp UTC`, "America/New_York")

ctd_macnorth %>% 
  select(`Datetime (EST)`, Depth) %>% 
  mutate(reef = "MacNorth") %>% 
  filter(`Datetime (EST)` >= as.Date("2022-12-06")) -> ctd_macnorth_depth
```

```{r}
rbind(ctd_emerald_depth, ctd_macnorth_depth) %>% 
  mutate(Season = case_when(`Datetime (EST)` < as.Date("2023-05-01") ~ "Dry",
                            `Datetime (EST)` >= as.Date("2023-05-01") ~ "Wet")) %>% 
  group_by(reef, Season) %>% 
  ggplot(., aes(x= `Datetime (EST)` , y=Depth, color = reef)) +
  theme_classic() +
  geom_point() +
  labs(x="Date", y = "Depth (m)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) + 
  theme(text = element_text(size = 15))
ggsave("depth_continuous.pdf")


rbind(ctd_emerald_depth, ctd_macnorth_depth) %>% 
  mutate(Season = case_when(`Datetime (EST)` < as.Date("2023-05-01") ~ "Dry",
                            `Datetime (EST)` >= as.Date("2023-05-01") ~ "Wet")) %>% 
  group_by(reef, Season) %>% 
  ggplot(., aes(x= Season , y=Depth, fill = reef)) +
  theme_classic() +
  geom_boxplot() +
  labs(x="Season", y = "Depth (m)") +
  scale_fill_manual(values = c("#edb9c3", "#5370c9")) + 
  theme(text = element_text(size = 15))
ggsave("depth_boxplot.pdf")

rbind(ctd_emerald_depth, ctd_macnorth_depth) -> depth_data

depth_plot<- ggplot(depth_data, aes(x= `Datetime (EST)` , y=Depth, color = reef)) +
  theme_classic() +
  geom_line() +
  labs(x="Date", y = "Depth (m)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) + 
  theme(text = element_text(size = 15)) +
  theme(legend.position = "none")
```


# pH

```{r}
library(openair)
library(seacarb)
```

import seafet data
```{r}
phDataRaw_emerald <- read_excel("raw_data/seafet-emerald-composite.xlsx", sheet = "seafet-emerald-composite")

phDataRaw_emerald$`Datetime (EST)` <- with_tz(phDataRaw_emerald$`DateTime (UTC+00:00)`, "America/New_York")

phDataRaw_emerald %>% 
  mutate(reef = "Emerald") %>% 
  select(!Swap) -> pH_emerald


phDataRaw_macnorth <- read_excel("raw_data/seafet-macnorth-composite.xlsx", sheet = "seafet-macnorth-composite")

phDataRaw_macnorth$`Datetime (EST)` <- with_tz(phDataRaw_macnorth$`DateTime (UTC+00:00)`, "America/New_York")

phDataRaw_macnorth %>% 
  mutate(reef = "MacNorth") -> pH_macnorth

rbind(pH_emerald, pH_macnorth) -> all_ph_data
```

round temperature data 
```{r}
str_macnorth$dateTimeEST5Rounded <- round_date(str_macnorth$`Datetime (EST)`, "5 minutes")

str_all %>% 
  full_join(ctd_emerald_temp_sept9) %>% 
  filter(reef=="Emerald") %>% 
  select(!Time) ->emerald_temp_data

emerald_temp_data$dateTimeEST5Rounded <- round_date(emerald_temp_data$`Datetime (EST)`, "5 minutes")
```


import bottle sample data
```{r}
bottle_data<-read_csv("raw_data/bottle-data.csv")

bottle_data$Date <- mdy(bottle_data$collected_date)
bottle_data$Time <- hms(bottle_data$collected_time)
bottle_data$rDateTime <-paste(bottle_data$Date, bottle_data$Time, sep = " ")
bottle_data$rDateTime <- ymd_hms(bottle_data$rDateTime)
bottle_data$rDateTime30Rounded <- round_date(bottle_data$rDateTime, "30 minutes")
bottle_data$dateTimeEST5Rounded <- round_date(bottle_data$rDateTime, "5 minutes")

bottle_data %>% 
  mutate(reef = case_when(collected_site == "Emerald Reef" ~ "Emerald",
                          collected_site == "Emerald Ree" ~ "Emerald",
                          collected_site == "MacArthur Cswy North" ~ "MacNorth",
                          collected_site == "MacArthur Causeway North" ~ "MacNorth",
                          collected_site == "Mac North" ~ "MacNorth")) -> bottle_data

bottleIter = nrow(bottle_data);#count number of rows in bottleDataRaw
#look up seafet pH (internal) and str temperature for each bottle, where present
for(i in 1:bottleIter){
    if (bottle_data$reef[i] == "Emerald") {#determine site
    phLookupLine = match(bottle_data$rDateTime30Rounded[i], pH_emerald$`Datetime (EST)`)#identify seafet row number for that site and time
    bottle_data$phInt[i] = pH_emerald$`Internal pH (pH)`[phLookupLine]#grab the pH from that line and put it into the bottle data table
    tempLookupLine = match(bottle_data$dateTimeEST5Rounded[i], str_emerald$dateTimeEST5Rounded)#identify the str row number for that site and time
    bottle_data$tempSTR[i] = emerald_temp_data$Temperature[tempLookupLine]#grab the str temperature from that line and put it into the bottle data table
  }
   else if (bottle_data$reef[i] == "MacNorth") {#determine site
    phLookupLine = match(bottle_data$rDateTime30Rounded[i], pH_macnorth$`Datetime (EST)`)#identify seafet row number for that site and time
    bottle_data$phInt[i] = pH_macnorth$`Internal pH (pH)`[phLookupLine]#grab the pH from that line and put it into the bottle data table
    tempLookupLine = match(bottle_data$dateTimeEST5Rounded[i], str_macnorth$dateTimeEST5Rounded)#identify the str row number for that site and time
    bottle_data$tempSTR[i] = str_macnorth$Temperature[tempLookupLine]#grab the str temperature from that line and put it into the bottle data table
   }
}
```


```{r}
#look at temp data comparison
bottle_data %>% 
  dplyr::select(collected_temp,rDateTime, reef, tempSTR) %>% 
  pivot_longer(c(collected_temp, tempSTR),names_to="temp_measure", values_to="temp_value") %>% 
  ggplot(., aes(x=rDateTime, y=temp_value, color=temp_measure)) +
  facet_wrap(~reef) +
  geom_point()

bottle_data %>% 
  dplyr::select(pH_measured,pH_calculated,rDateTime, reef, phInt) %>% 
  pivot_longer(c(pH_measured,pH_calculated,phInt),names_to="pH_measure", values_to="pH_value") %>% 
  ggplot(., aes(x=rDateTime, y=pH_value, color=pH_measure)) +
  facet_wrap(~reef) +
  geom_point()
#ggsave("bottlesamples_vs_seafet.pdf")
#definitely some differences between bottle samples and seafet measurements
```

```{r}
for(i in 1:bottleIter){
  if(is.na(bottle_data$collected_temp[i])){output = NA} #if tempSTR is absent, set output to NA
  else { 
    output = carb(flag=15, var1=bottle_data$TA_umol_kg[i]/1000000, var2=bottle_data$DIC_umol_kg[i]/1000000, S=bottle_data$calc_salinity[i],T=bottle_data$collected_temp[i], P=bottle_data$collected_depth[i], k1k2 = "l", kf = "pf", ks = "d", pHscale = "T")
    bottle_data$phCalc[i] = output[6]
  #calculates dic from pH and 
    outputDic = carb(flag=8, var1=bottle_data$pH_measured[i], var2=bottle_data$TA_umol_kg[i]/1000000,  S=bottle_data$calc_salinity[i],T=bottle_data$collected_temp[i], P=bottle_data$collected_depth[i], k1k2 = "l", kf = "pf", ks = "d", pHscale = "T")
    bottle_data$dicCalc[i] = outputDic[16]
    bottle_data$dicCalc[i] = as.numeric(bottle_data$dicCalc[i])
    bottle_data$pco2[i] = outputDic[13]
    bottle_data$omegaArag[i] = outputDic[18]
    }
}
```



```{r}
bottle_data
#create dic check
bottle_data$dicCalc<-(as.numeric(bottle_data$dicCalc)*1000000)
bottle_data$dicDiff = (bottle_data$DIC_umol_kg-bottle_data$dicCalc)
#make graphs for salinity normalization of TA and DIC

taEndPointGraph <- ggplot(bottle_data, aes(x = calc_salinity, y = TA_umol_kg)) +
  theme_light()+
  geom_point(aes(color = reef), size = 3)+
  scale_color_manual(values=c("#D69EA0","#2A5676"))+  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  geom_smooth(method = 'lm', formula = y~x)+
  stat_cor(label.x = 31, label.y = 2480) +
  stat_regline_equation(label.x = 31, label.y = 2500)
dicEndPointGraph <- ggplot(bottle_data, aes(x = calc_salinity, y = dicCalc)) +
  theme_light()+
  geom_point(aes(color = reef), size = 3)+
  scale_color_manual(values=c("#D69EA0","#2A5676"))+  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  geom_smooth(method = 'lm', formula = y~x)+
  stat_cor(label.x = 31, label.y = 2380) +
  stat_regline_equation(label.x = 31, label.y = 2400)
taDicPointGraph <- ggplot(bottle_data, aes(x = TA_umol_kg, y = DIC_umol_kg)) +
  theme_light()+
  geom_point(aes(color = reef), size = 3)+
  scale_color_manual(values=c("#D69EA0","#2A5676"))+  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  geom_smooth(method = 'lm', formula = y~x)+
  stat_cor(label.x = 2300, label.y = 2380) +
  stat_regline_equation(label.x = 2300, label.y = 2400)
taDicCalcPointGraph <- ggplot(bottle_data, aes(x = TA_umol_kg, y = dicCalc)) +
  theme_light()+
  geom_point(aes(color = reef), size = 3)+
  scale_color_manual(values=c("#D69EA0","#2A5676"))+  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  geom_smooth(method = 'lm', formula = y~x)+
  stat_cor(label.x = 2300, label.y = 2380) +
  stat_regline_equation(label.x = 2300, label.y = 2400)
dicDiffTimeGraph <- ggplot(bottle_data, aes(x = rDateTime, y = dicDiff)) +
  theme_light()+
  geom_point(aes(color = reef), size = 3)+
  scale_color_manual(values=c("#D69EA0","#2A5676"))+  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
#dic Issues Graph
ggarrange(taDicPointGraph, taDicCalcPointGraph, dicDiffTimeGraph, ncol = 3, nrow = 1)
ggarrange(taEndPointGraph,dicEndPointGraph, ncol = 2, nrow = 1)
#proceed with Friss salinity normalization from end point graph equations
taEndMember = 2200
dicEndMember = 3300
bottle_data$ta35Friss = 35*((bottle_data$TA_umol_kg-taEndMember)/bottle_data$calc_salinity)+taEndMember
bottle_data$dic35Friss = 35*((bottle_data$dicCalc-dicEndMember)/bottle_data$calc_salinity)+dicEndMember
#graph TA vs DIC
taDicFrissPointGraph <- ggplot(bottle_data, aes(x = dic35Friss, y = ta35Friss)) +
  theme_light()+
  geom_point(aes(color = reef), size = 3)+
  scale_color_manual(values=c("#D69EA0","#2A5676"))+  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  geom_smooth(method = 'lm', formula = y~x)+
  stat_cor(label.x = 2000, label.y = 2500) +
  stat_regline_equation(label.x = 2000, label.y = 2520)
#create ta dic graph
taDicFrissPointGraph
#create ta and dc vs salinity graphs
ggarrange(taEndPointGraph,dicEndPointGraph, taDicFrissPointGraph,ncol = 3, nrow = 1)
#ggsave("TA_DIC_diagnosticgraphs.pdf", width = 10, height=7)
```


```{r}
library(plyr)
library(dplyr)

#create bottle tables for all samples
bottle_data$pco2<-as.numeric(bottle_data$pco2)
bottle_data$omegaArag<-as.numeric(bottle_data$omegaArag)

#add season
numbRows = nrow(bottle_data)
for(i in 1:numbRows){
  if (bottle_data$rDateTime[i]>"2023-5-26" && bottle_data$rDateTime[i]<"2023-10-15"){bottle_data$precipitation[i] = "wet"}
  else if (bottle_data$rDateTime[i]>"2022-5-26" && bottle_data$rDateTime[i]<"2022-10-15"){bottle_data$precipitation[i] = "wet"}
  else {bottle_data$precipitation[i] = "dry"}
}


bottleSummary <- ddply(bottle_data, c("reef","precipitation"), summarise,
                             countSalinityDensitometer = length(dicCalc),
                             meanSalinityDensitometer = round(mean(calc_salinity),2),
                             sdSalinityDensitometer   = round(sd(calc_salinity),2),
                             meanSpecPhField = round(mean(pH_measured, na.rm=TRUE),3),
                             sdSpecPhField   = round(sd(pH_measured, na.rm=TRUE),3),
                             meanTaCorrected = round(mean(TA_umol_kg, na.rm=TRUE),1),
                             sdTaCorrected   = round(sd(TA_umol_kg, na.rm=TRUE),1),
                             meanDicCalc = round(mean(dicCalc, na.rm=TRUE),1),
                             sdDicCalc   = round(sd(dicCalc, na.rm=TRUE),2),
                             meanTaFriss = round(mean(ta35Friss, na.rm=TRUE),1),
                             sdTaFriss   = round(sd(ta35Friss, na.rm=TRUE),1),
                             meanDicFriss = round(mean(dic35Friss, na.rm=TRUE),1),
                             sdDicFriss   = round(sd(dic35Friss, na.rm=TRUE),1),                            
                             meanPco2 = round(mean(pco2, na.rm=TRUE),0),
                             sdPco2   = round(sd(pco2, na.rm=TRUE),0),
                             meanOmegaArag = round(mean(omegaArag, na.rm=TRUE),2),
                             sdOmegaArag   = round(sd(omegaArag, na.rm=TRUE),2)
)
bottleSummary
```


```{r}
#plot TA and DIC as a function of time
dicTimeGraph <- ggplot(bottle_data, aes(x = collected_time, y = dicCalc)) +
  theme_light()+
  geom_point(aes(color = reef, shape = precipitation), size = 3)+
  scale_color_manual(values=c("#D69EA0","#2A5676")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
taTimeGraph <- ggplot(bottle_data, aes(x = collected_time, y = TA_umol_kg)) +
  theme_light()+
  geom_point(aes(color = reef, shape = precipitation), size = 3)+
  scale_color_manual(values=c("#D69EA0","#2A5676"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

ggarrange(taTimeGraph,dicTimeGraph, ncol = 2, nrow = 1)
#ggsave("TA_DIC_timeofday.pdf",width=10,height=7)
```

```{r}
#calculate the particular seafet offset for each bottle sample
for(i in 1:bottleIter){
  if(is.na(bottle_data$pH_measured[i])){#if spec pH field is absent, use the pH calculated from dic and ta
#    bottleDataRaw$offset[i] =1
    bottle_data$offset[i] = (bottle_data$phInt[i] - as.numeric(bottle_data$phCalc[i]))
    bottle_data$phUsedOffset[i] = as.numeric(bottle_data$phCalc[i])
    bottle_data$offsetSource[i] = "calc"
    
  } 
  else {
#    bottleDataRaw$offset[i] =2
    bottle_data$offset[i] = (bottle_data$phInt[i] - bottle_data$pH_measured[i])
    bottle_data$phUsedOffset[i] = bottle_data$pH_measured[i]
    bottle_data$offsetSource[i] = "spec"
  }
}
```

```{r}
pH_plot<-ggplot()+
  geom_line(data = all_ph_data, aes(x = `Datetime (EST)`, y = `Internal pH (pH)`,  color = reef)) +
  geom_point(data = bottle_data, aes(x = rDateTime, y = phUsedOffset, color = reef, shape = offsetSource)) +
  scale_shape_manual(values=c(15, 19))+
  scale_color_manual(values=c("#edb9c3", "#5370c9"))+  
  theme_classic()+
  scale_y_continuous(name = "pH Int", limits = c(7.15,8.15), labels = scales::number_format(accuracy = 0.1))+
  theme(text = element_text(size = 15))+
  geom_line() +
  theme(legend.position = "none")
#ggsave("seafetdata_uncorrected.pdf")
```

use offset to adjust internal pH values from seafets
```{r}
#first break up year-long dataset into collection time points so offsets from each deployment match

bottle_data %>% 
  select(collected_date,reef,offset) %>% 
  drop_na() %>% 
  mutate(collected_date=mdy(collected_date)) -> offsets

phIter = nrow(all_ph_data);#count number of rows in bottleDataRaw
#look up seafet pH (internal) and str temperature for each bottle, where present
for(i in 1:phIter){
  if ((all_ph_data$reef[i] == "Emerald") && (all_ph_data$`Datetime (EST)`[i] < as.Date("2023-03-28"))) {all_ph_data$phFinal[i]=all_ph_data$`Internal pH (pH)`[i]-offsets[4,3]}
  else if ((all_ph_data$reef[i] == "MacNorth") && (all_ph_data$`Datetime (EST)`[i] < as.Date("2023-03-28"))) {all_ph_data$phFinal[i]=all_ph_data$`Internal pH (pH)`[i]-offsets[5,3]}
  else if ((all_ph_data$reef[i] == "Emerald") && (all_ph_data$`Datetime (EST)`[i] < as.Date("2023-07-11"))) {all_ph_data$phFinal[i]=all_ph_data$`Internal pH (pH)`[i]-offsets[6,3]}
  else if ((all_ph_data$reef[i] == "MacNorth") && (all_ph_data$`Datetime (EST)`[i] < as.Date("2023-07-11"))) {all_ph_data$phFinal[i]=all_ph_data$`Internal pH (pH)`[i]-offsets[7,3]}
  else if ((all_ph_data$reef[i] == "Emerald") && (all_ph_data$`Datetime (EST)`[i] < as.Date("2023-11-20"))) {all_ph_data$phFinal[i]=all_ph_data$`Internal pH (pH)`[i]-offsets[8,3]}
  else if ((all_ph_data$reef[i] == "MacNorth") && (all_ph_data$`Datetime (EST)`[i] < as.Date("2023-11-20"))) {all_ph_data$phFinal[i]=all_ph_data$`Internal pH (pH)`[i]-offsets[9,3]}
  else{all_ph_data$phFinal[i]=NA}
}

all_ph_data$final_pH <- as.numeric(all_ph_data$phFinal)

ggplot()+
  geom_line(data = all_ph_data, aes(x = `Datetime (EST)`, y = final_pH,  color = reef)) +
  geom_point(data = bottle_data, aes(x = rDateTime, y = phUsedOffset, color = reef, shape = offsetSource)) +
  scale_shape_manual(values=c(15, 19))+
  scale_color_manual(values=c("#D69EA0","#2A5676"))+  
  theme_classic()+
  theme(plot.title = element_text(size = 15))+
  geom_line()+
  theme(legend.position = "none")
#ggsave("seafet_correctedwithbottlesampleoffset.pdf")
```

plot these together
```{r}
cowplot::plot_grid(parplot,depth_plot,pH_plot, ncol=3)
ggsave("par_depth_pH.pdf",width=15,height=7)
```




# Tilt

load packages
```{r}
library(RColorBrewer)
library(scales)
library(lubridate)
library(openair)
```

```{r}
tilt_emerald<- read_xlsx("tilt-emerald-sn2102069-redacted.xlsx", sheet = "tilt-emerald-sn2102069")
tilt_macnorth<- read_xlsx("tilt-macnorth-sn2102068-redacted.xlsx", sheet = "tilt-macnorth-sn2102068")

tilt_emerald$DateTime_EST <- with_tz(tilt_emerald$DateTime, "America/New_York")
tilt_macnorth$DateTime_EST <- with_tz(tilt_macnorth$DateTime, "America/New_York")

tilt_emerald %>% 
  rename(ws = `Speed (cm/s)`, wd = `Heading (degrees)`) %>% 
  filter(DateTime_EST >= as.Date("2022-12-06")) %>% 
  mutate(reef = "Emerald") -> tilt_emerald_clean

tilt_macnorth %>% 
  rename(ws = `Speed (cm/s)`, wd = `Heading (degrees)`) %>% 
  filter(DateTime_EST >= as.Date("2022-12-06")) %>% 
  mutate(reef = "MacNorth") -> tilt_macnorth_clean

windRose(tilt_emerald_clean, ws.int=3, max.freq = 60, cols = brewer_pal(palette="Spectral", direction=-1)(6), breaks=6, paddle=FALSE, key.header="(cm/s)", key.footer="Emerald Reef")
ggsave("tilt_emerald.png")


windRose(tilt_macnorth_clean, ws.int=6, max.freq = 60, cols = brewer_pal(palette="Spectral", direction=-1)(6), breaks=6, paddle=FALSE, key.header="(cm/s)", key.footer="MacArthur North")
ggsave("tilt_macnorth.pdf")
```


# old data analysis 

## temp
## calculating maximum monthly mean temperature over experiment time period for each reef
```{r}
str_all %>% 
  filter(Date >= "2022-12-06") %>% 
  full_join(ctd_emerald_temp_sept9) -> temp_data
 
temp_data %>% 
  select(!Date) %>% 
  separate(`Datetime (EST)`, into = c("Date", "Time"), sep = " ") %>% 
  mutate(Date = ymd(Date)) %>% 
  group_by(Date, reef) %>% 
  summarise(max_temp = max(Temperature)) %>% 
  mutate(year_month = paste(year(Date), month(Date), sep = "-")) %>% 
  mutate(year_month = ym(year_month)) %>% 
  group_by(year_month, reef) %>% 
  summarise(mean_max_temp = mean(max_temp)) %>% 
  group_by(reef) %>% 
  summarise(MMM = mean(mean_max_temp))
```

## calculate the maximum summertime mean
```{r}
temp_data %>% 
  select(!Date) %>% 
  filter(as.Date(`Datetime (EST)`) >= "2023-06-01" & as.Date(`Datetime (EST)`) <= "2023-09-01") %>% 
  separate(`Datetime (EST)`, into = c("Date", "Time"), sep = " ") %>% 
  mutate(Date = ymd(Date)) %>% 
  group_by(Date, reef) %>% 
  summarise(max_temp = max(Temperature)) %>% 
  mutate(year_month = paste(year(Date), month(Date), sep = "-")) %>% 
  mutate(year_month = ym(year_month)) %>% 
  group_by(year_month, reef) %>% 
  summarise(mean_max_temp = mean(max_temp)) %>% 
  group_by(reef) %>% 
  summarise(MMM = mean(mean_max_temp)) #mean summertime temp
```


## calculate DHW (weeks spent above the mean maximum summertime temp)
```{r}
temp_data %>% 
  filter(reef == "Emerald") %>% 
  filter(Temperature > 30.16728) %>% 
  distinct(Date) #89 days above?? That is 12.7 weeks

temp_data %>% 
  filter(reef == "Emerald") %>% 
  filter(Temperature > 31) 
```

## graph focusing on summertime
```{r}
str_all %>% 
  full_join(ctd_emerald_temp_sept9) %>% 
  filter(as.Date(`Datetime (EST)`) <= "2023-10-01" & as.Date(`Datetime (EST)`) >= "2023-05-31") %>% 
ggplot(., aes(x=`Datetime (EST)`, y=Temperature, color = reef)) +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "Temperature (ºC)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) +
  geom_hline(yintercept = 31, linetype = "dashed", color = "black") +
  annotate("text", x = as.POSIXct("2023-10-01"), y = 31, label = "31ºC", vjust = -1) +
  annotate("text", x = as.POSIXct("2023-09-30"), y = 33, label= "Days Above 31ºC:\nEmerald=44\nMacNorth=85", colour = "black", hjust=1) 

# Display the plot

temp_data %>% 
  select(!Date) %>% 
  filter(as.Date(`Datetime (EST)`) >= "2023-06-01" & as.Date(`Datetime (EST)`) <= "2023-10-01") %>% 
  separate(`Datetime (EST)`, into = c("Date", "Time"), sep = " ") %>% 
  mutate(Date = ymd(Date)) %>% 
  group_by(Date, reef) %>% 
  summarise(max_temp = max(Temperature)) %>% 
  filter(max_temp > 31) %>% 
  group_by(reef) %>% 
  summarise(count = n())
```

## daily variability experienced
```{r}
temp_data %>% 
  select(!Date) %>% 
  separate(`Datetime (EST)`, into = c("Date", "Time"), sep = " ") %>% 
  mutate(Date = ymd(Date)) %>% 
  group_by(Date, reef) %>% 
  summarise(daily_min = min(Temperature), daily_max=max(Temperature)) %>% 
  mutate(daily_range=daily_max - daily_min) %>% 
  ungroup() %>% 
  ggplot(., aes(x=Date, y=daily_range, color = reef)) +
  geom_point() +
  geom_line() +
  labs(x="Day", y = "Daily range (ºC)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) + 
  theme_classic()

temp_data %>% 
  select(!Date) %>% 
  separate(`Datetime (EST)`, into = c("Date", "Time"), sep = " ") %>% 
  mutate(Date = ymd(Date)) %>% 
  group_by(Date, reef) %>% 
  summarise(daily_min = min(Temperature), daily_max=max(Temperature)) %>% 
  mutate(daily_range=daily_max - daily_min) %>% 
  ungroup() %>% 
  ggplot(., aes(x=Date, y=daily_range, color = reef)) +
  geom_point() +
  geom_smooth() +
  labs(x="Day", y = "Daily range (ºC)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) + 
  theme_classic() +
  theme(text = element_text(size = 15))
#ggsave("dailyrangetemp.pdf")
```


## Separating by season and summarizing temp data
```{r}
temp_data %>% 
  select(!Date) %>% 
  separate(`Datetime (EST)`, into = c("Date", "Time"), sep = " ") %>% 
  mutate(Date = ymd(Date)) %>% 
  mutate(Season = case_when(Date < "2023-05-01" ~ "Dry",
                            Date >= "2023-05-01" ~ "Wet")) %>% 
  ggplot(., aes(x=Season, y=Temperature, fill = reef)) +
  geom_boxplot() +
  labs(x="Season", y = "Temp (ºC)") +
  scale_fill_manual(values = c("#edb9c3", "#5370c9")) + 
  theme_classic()
#ggsave("boxplot_temp_byseason.pdf")
```


```{r}
temp_data %>% 
  select(!Date) %>% 
  separate(`Datetime (EST)`, into = c("Date", "Time"), sep = " ") %>% 
  mutate(Date = ymd(Date)) %>% 
  mutate(Season = case_when(Date < "2023-05-01" ~ "Dry",
                            Date >= "2023-05-01" ~ "Wet")) %>% 
  group_by(Season, reef) %>% 
  summarise(count=n(), mean_temp = mean(Temperature), sd_temp = sd(Temperature), min_temp = min(Temperature), max_temp= max(Temperature), range_temp = max_temp-min_temp) %>% 
  write_csv("season_reef_tempsummary.csv")
```



## EcoPAR

### ecopar1959 raw txt data file
```{r}
#look at raw data file 
ecopar1959_032823 <- read_tsv("environmental_data/nov22-mar23_rawdata/ecopar-sn5954-tag1959.raw.txt")

ecopar1959_032823 %>% 
  separate(col = `53651 records to read`, into = c("Date", "Time", "Value"), sep = "\t") %>% 
  mutate(Date = mdy(Date), Time = hms(Time), Value = as.numeric(Value)) -> ecopar1959_032823

length(unique(ecopar1959_032823$Date)) #115 days

ecopar1959_032823 %>% 
  filter(Date == "2022-12-31") #10 measurements were taken instantaneously at every 30 min interval for each day.

ecopar1959_032823$datetime <- paste(ecopar1959_032823$Date, ecopar1959_032823$Time)
ecopar1959_032823$datetime <- ymd_hms(ecopar1959_032823$datetime)
  
ecopar1959_032823 %>% #53,652 values
  mutate(interval = floor_date(datetime, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(Value)) %>% #5,147 values
  ggplot(., aes(x=interval, y=mean_value)) + geom_point()


# look at Mike's data tidied excel file (it looks like he made a formula to convert PAR(raw) to PAR(calc). Formula is [Digital Im] * Power(10, PARraw - [Digital a0/Digitial a1])). Unsure what the digitial things are but it changes the PAR value by approximately 10^3 (6000 -> 6)

#I found the equations from the SeaBird Scientific ECO Photosynthetically Active Radiation (PAR) sensor documentation: https://www.seabird.com/eco-photosynthetically-active-radiation-par-sensor/product-downloads?id=60762467733

#Underwater PAR
#Seasoft calculates light from an underwater Satlantic logarithmic analog PAR sensor as:
# PAR = multiplier * Im * 10 (V –a0) / a1 [user-selected units]
# where
# V = voltage output
# a0, a1, and lm are from the PAR calibration sheet that was provided by Satlantic
# multiplier = 1.0 for units of μmol photons/ m2 sec (see Notes)

#so the unit is micromol. convert to mol/day to be able to compare with Ian's paper
```



### ecopar1959 (Emerald) Dec-Mar Excel tidy data
```{r}
ecopar1959_032823_excel <-read_xlsx("Dec2022-Mar2023_rough-ecopar-sn5954-tag1959-emerald.xlsx", sheet = "sheet1")

ecopar1959_032823_excel$`Datetime (UTC)` <- as.POSIXct(ecopar1959_032823_excel$`Datetime (UTC)`, tz = "UTC")
ecopar1959_032823_excel$`Datetime (EST)` <- with_tz(ecopar1959_032823_excel$`Datetime (UTC)`, "America/New_York")

ecopar1959_032823_excel %>% #53,621 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  ggplot(., aes(x=interval, y=mean_value)) + geom_point() #values look a lot different and cleaner. I'll use these for plotting

# in Enochs et al. 2023, they "summed [PAR readings] between the hours of 10am and 3pm, to calculate a daily dose."

ecopar1959_032823_excel %>% #53,621 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  group_by(as.Date(`Datetime (EST)`)) %>% 
  mutate(hour = hour(`Datetime (EST)`), Date =as.Date(`Datetime (EST)`) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(`PAR (calc)`, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(`PAR (calc)`)) %>% 
  ggplot(., aes(x= Date , y=daily_dose)) +
  geom_point() +
  geom_line() +
  theme_classic() # these values look way higher than what Ian published as the daily dose, the range of his graph is 0-30 mol/day. I wonder if he took the mean value for the 10 instantaneous values taken at each 30 min? because adding those all together isn't an accurate representation.

ecopar1959_032823_excel %>% #53,621 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  ggplot(., aes(x= Date , y=daily_dose)) +
  geom_point() +
  geom_line() +
  theme_classic() #well that just reduced everything by an order of 10 but otherwise still way too high (~6000). That is because this is micromol. Divide by 1000 to get mol/day


ecopar1959_032823_excel %>% #53,621 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  ggplot(., aes(x= Date , y=PAR_mol_day)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "PAR (mol/day)")
```

### ecopar4298 (Emerald) Apr-Jul Excel tidy data
```{r}
ecopar4298_071123_excel <- read_xlsx("apr23-jul23_rawdata/ecopar-sn5954-tag4298_Emerald.xlsx", sheet = "ecopar-sn5954-tag4298.raw")

ecopar4298_071123_excel$`Datetime (UTC)` <- as.POSIXct(ecopar4298_071123_excel$`Datetime (UTC)`, tz = "UTC")
ecopar4298_071123_excel$`Datetime (EST)` <- with_tz(ecopar4298_071123_excel$`Datetime (UTC)`, "America/New_York")

ecopar4298_071123_excel %>% #47,040 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  ggplot(., aes(x= Date , y=PAR_mol_day)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "PAR (mol/day)")
```


### ecopar1960 (MacN) Dec-Mar Excel tidy data
```{r}
ecopar1960_032823_excel <-read_excel("Dec2022-Mar2023_rough-ecopar-sn5953-tag1960-macnorth.xlsx", sheet = "ecopar-sn5953-tag1960.raw")

ecopar1960_032823_excel$`Datetime (UTC)` <- as.POSIXct(ecopar1960_032823_excel$`Datetime (UTC)`, tz = "UTC")
ecopar1960_032823_excel$`Datetime (EST)` <- with_tz(ecopar1960_032823_excel$`Datetime (UTC)`, "America/New_York")

ecopar1960_032823_excel %>% #47,040 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  ggplot(., aes(x= Date , y=PAR_mol_day)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "PAR (mol/day)")
```

### ecopar4297 (MacN) Apr-Jul Excel tidy data
```{r}
ecopar4297_071123_excel <- read_excel("apr23-jul23_rawdata/ecopar-sn5953-tag4297_MacNorth.xlsx", sheet = "ecopar-sn5953-tag4297.raw")

ecopar4297_071123_excel$`Datetime (UTC)` <- as.POSIXct(ecopar4297_071123_excel$`Datetime (UTC)`, tz = "UTC")
ecopar4297_071123_excel$`Datetime (EST)` <- with_tz(ecopar4297_071123_excel$`Datetime (UTC)`, "America/New_York")

ecopar4297_071123_excel %>% #47,040 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  ggplot(., aes(x= Date , y=PAR_mol_day)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "PAR (mol/day)")
```

### combine all datasets into 1 so they can all be plotted together

```{r}
ecopar1959_032823_excel %>% #53,621 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  mutate(reef = "Emerald") -> ecopar1959_032823

ecopar4298_071123_excel %>% #47,040 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  mutate(reef = "Emerald") -> ecopar4298_071123 

ecopar1960_032823_excel %>% #47,040 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  mutate(reef = "MacNorth") -> ecopar1960_032823

ecopar4297_071123_excel %>% #47,040 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  mutate(reef = "MacNorth") -> ecopar4297_071123

full_join(ecopar1959_032823, ecopar4298_071123) %>% 
  full_join(., ecopar1960_032823) %>% 
  full_join(., ecopar4297_071123) -> all_ecopar_data_molperday

all_ecopar_data_molperday %>% 
  ggplot(., aes(x= Date , y=PAR_mol_day, color = reef)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "PAR (mol/day)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9"))
```

### Stats 

```{r}
all_ecopar_data_molperday

#group dates by season
#wet season = mid-may -- mid-october
#dry season = mid-october -- mid-may

all_ecopar_data_molperday %>% 
  mutate(season = case_when(Date >= "2023-05-15" ~ "Wet",
                            Date < "2023-05-15" ~ "Dry")) -> all_ecopar_data_molperday

ecopar_lm <- lm(PAR_mol_day ~ reef*season, data = all_ecopar_data_molperday)

summary(ecopar_lm)

ecopar_anova<- aov(ecopar_lm)
summary(ecopar_anova)

qqnorm(ecopar_anova$residuals) #assume normality
boxplot(resid(ecopar_anova)~all_ecopar_data_molperday$reef)
boxplot(resid(ecopar_anova)~all_ecopar_data_molperday$season) #homogeneity of variance

TukeyHSD(ecopar_anova)


```



## SeaFET



## CTD

### CTD6635 (Emerald) Dec-Mar Excel tidy data
```{r}
ctd6635_032823 <- read_xlsx("Dec2022-Mar2023_rough-ctd-emerald-6635.xlsx", sheet = "emerald-ctd-6635-data-edit")

#plot time values as is and see if the trend for time of day is weird (then it might be UTC not EST)
ctd6635_032823 %>% 
  select(Timestamp:Depth, Salinity) %>% 
  mutate(Timestamp = dmy_hms(Timestamp)) %>% 
  mutate(reef = "Emerald") -> ctd6635_032823
```

### CTD7398 (Emerald) Apr-Jul Excel data
```{r}
ctd7398_071123 <- read_xlsx("apr23-jul23_rawdata/ctd-sn7398-edited.xlsx", sheet = "ctd-sn7398-edited")

ctd7398_071123 %>% 
  select(Timestamp.UTC, Temp, Cond, Depth, Salinity) %>% 
  rename(Timestamp = Timestamp.UTC, Temperature = Temp, Conductivity = Cond) %>% 
  filter(as.Date(Timestamp) > "2023-03-28") %>% 
  mutate(reef = "Emerald") -> ctd7398_071123
```

### CTD7398 (MacN) Dec-Mar Excel
```{r}
ctd7398_032823 <- read_xlsx("Dec2022-Mar2023_rough-ctd-macnorth-7398.xlsx", sheet = "macnorth-ctd-7398-data-edit")

ctd7398_032823 %>% 
  select(Timestamp, Temperature, Conductivity, Depth, Salinity) %>% 
  mutate(Timestamp = dmy_hms(Timestamp)) %>% 
  mutate(reef = "MacNorth") -> ctd7398_032823
```

### CTD6635 (MacN) Apr-Jul 
```{r}
#data not tidied yet

#try to read the hex file
library(hexView)

hexViewFile("~/Downloads/SBE16plus_01606635_2023_07_13.hex")

#that didn't work it's not readable. needs to be run through the CTD software to extract readable numbers
```


### combine files and build temperature graph
```{r}
full_join(ctd6635_032823, ctd7398_071123) %>% 
  full_join(., ctd7398_032823) %>% 
  filter(as.Date(Timestamp) < "2023-04-01") %>% 
  ggplot(., aes(x=Timestamp, y=Temperature, color = reef)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "Temperature (ºC)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9"))
```
### combine files and build salinity graph
```{r}
full_join(ctd6635_032823, ctd7398_071123) %>% 
  full_join(., ctd7398_032823) %>% 
  filter(as.Date(Timestamp) < "2023-04-01") %>% 
  filter(Salinity > 31) %>% 
  ggplot(., aes(x=Timestamp, y=Salinity, color = reef)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "Salinity") +
  scale_color_manual(values = c("#edb9c3", "#5370c9"))
```

