---
title: "environmental_data_plots"
author: "allyson_demerlis"
date: "2023-08-31"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
library(cowplot)
```

Mike Jankulak created composite files of all time points for each site for each instrument. Some instruments were swapped between sites, so he made sure to group the sites correctly. There is a metadata file that shows all the instrument deployments.

# Temperature

## import STR files
```{r}
str_emerald <- read_excel("str-emerald-sn5042.xlsx", sheet = "str-emerald-sn5042")

str_emerald$`Datetime (UTC)` <- as.POSIXct(str_emerald$`Datetime (UTC)`, tz="UTC")
str_emerald$`Datetime (EST)` <- with_tz(str_emerald$`Datetime (UTC)`, "America/New_York")

str_emerald %>% 
  mutate(reef = "Emerald") %>% 
  select(!Time) %>% 
  select(!`Datetime (UTC)`) %>% 
  select(!`Sample Number`) -> str_emerald
  
str_macnorth <- read_excel("str-macnorth-sn9235.xlsx", sheet = "str-macnorth-sn9235")

str_macnorth$Datetime <- as.POSIXct(str_macnorth$Datetime, tz="UTC")
str_macnorth$`Datetime (EST)` <- with_tz(str_macnorth$Datetime, "America/New_York")

str_macnorth %>% 
  mutate(reef = "MacNorth") %>% 
  select(!Datetime) %>% 
  select(!`Sample Number`) %>% 
  select(!Time) -> str_macnorth
```


## just STR data
```{r}
full_join(str_emerald, str_macnorth) -> str_all

# Identify the lowest and highest points
lowest_point <- str_all %>% filter(Temperature == min(Temperature))
highest_point <- str_all %>% filter(Temperature == max(Temperature))

str_all %>% filter(reef == "Emerald") %>% filter(Temperature == max(Temperature))

str_all %>% 
  filter(Date >= "2022-12-06") %>% 
ggplot(., aes(x=`Datetime (EST)`, y=Temperature, color = reef)) +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "Temperature (ºC)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) +
  geom_hline(yintercept = 31, linetype = "dashed", color = "black") +
  annotate("text", x = as.POSIXct("2023-11-20"), y = 31, label = "31ºC", vjust = -1) +
  geom_point(data = lowest_point, aes(`Datetime (EST)`, Temperature), color = "#5370c9", size = 2) +
  geom_text(data = lowest_point, aes(`Datetime (EST)`, Temperature, label = "17.89ºC"), vjust = 1.1, hjust = -0.1, color = "black") +
  geom_point(data = highest_point, aes(`Datetime (EST)`, Temperature), color = "#5370c9", size = 2) +
  geom_text(data = highest_point, aes(`Datetime (EST)`, Temperature, label = "33.55ºC"), vjust = 0, hjust = -0.2, color = "black") +
labs(title = "2022-2023 Temperature Data at Emerald Reef and MacArthur Causeway", x="Month") 
  #annotate("text", x = as.POSIXct("2023-09-30"), y = 18, label= "STR data at Emerald stopped 9/9", colour = "black", hjust = .56)
```

## append CTD temperature data for Emerald after 9/9/23
```{r}
ctd_emerald <- read_excel("ctd-emerald-composite.xlsx", sheet = "ctd-emerald-composite")

ctd_emerald %>% 
  select(`Timestamp UTC`, Temperature) -> ctd_emerald_temp

ctd_emerald_temp$`Timestamp (UTC)` <- as.POSIXct(ctd_emerald_temp$`Timestamp UTC`, tz="UTC")
ctd_emerald_temp$`Datetime (EST)` <- with_tz(ctd_emerald_temp$`Timestamp (UTC)`, "America/New_York")

ctd_emerald_temp %>% 
  mutate(Date = as.Date(`Datetime (EST)`)) %>% #the dates towards midnight don't match Datetime, it rounds up to the next day
  select(-1,-3) %>% 
  mutate(reef = "Emerald") %>% 
  filter(Date >= "2023-09-09") -> ctd_emerald_temp_sept9

str_all %>% 
  filter(Date >= "2022-12-06") %>% 
  full_join(ctd_emerald_temp_sept9) %>% 
ggplot(., aes(x=`Datetime (EST)`, y=Temperature, color = reef)) +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "Temperature (ºC)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) +
  geom_hline(yintercept = 31, linetype = "dashed", color = "black") +
  annotate("text", x = as.POSIXct("2023-11-20"), y = 31, label = "31ºC", vjust = -1) +
  geom_point(data = lowest_point, aes(`Datetime (EST)`, Temperature), color = "#5370c9", size = 2) +
  geom_text(data = lowest_point, aes(`Datetime (EST)`, Temperature, label = "17.89ºC"), vjust = 1.1, hjust = -0.1, color = "black") +
  geom_point(data = highest_point, aes(`Datetime (EST)`, Temperature), color = "#5370c9", size = 2) +
  geom_text(data = highest_point, aes(`Datetime (EST)`, Temperature, label = "33.55ºC"), vjust = 0, hjust = -0.2, color = "black") +
labs(title = "2022-2023 Temperature Data at Emerald Reef and MacArthur Causeway", x="Month") 
```

## calculating maximum monthly mean temperature over experiment time period for each reef
```{r}
str_all %>% 
  filter(Date >= "2022-12-06") %>% 
  full_join(ctd_emerald_temp_sept9) -> temp_data
 
temp_data %>% 
  select(!Date) %>% 
  separate(`Datetime (EST)`, into = c("Date", "Time"), sep = " ") %>% 
  mutate(Date = ymd(Date)) %>% 
  group_by(Date, reef) %>% 
  summarise(max_temp = max(Temperature)) %>% 
  mutate(year_month = paste(year(Date), month(Date), sep = "-")) %>% 
  mutate(year_month = ym(year_month)) %>% 
  group_by(year_month, reef) %>% 
  summarise(mean_max_temp = mean(max_temp)) %>% 
  group_by(reef) %>% 
  summarise(MMM = mean(mean_max_temp))
```

## calculate the maximum summertime mean
```{r}
temp_data %>% 
  select(!Date) %>% 
  filter(as.Date(`Datetime (EST)`) >= "2023-06-01" & as.Date(`Datetime (EST)`) <= "2023-09-01") %>% 
  separate(`Datetime (EST)`, into = c("Date", "Time"), sep = " ") %>% 
  mutate(Date = ymd(Date)) %>% 
  group_by(Date, reef) %>% 
  summarise(max_temp = max(Temperature)) %>% 
  mutate(year_month = paste(year(Date), month(Date), sep = "-")) %>% 
  mutate(year_month = ym(year_month)) %>% 
  group_by(year_month, reef) %>% 
  summarise(mean_max_temp = mean(max_temp)) %>% 
  group_by(reef) %>% 
  summarise(MMM = mean(mean_max_temp)) #mean summertime temp
```


## calculate DHW (weeks spent above the mean maximum summertime temp)
```{r}
temp_data %>% 
  filter(reef == "Emerald") %>% 
  filter(Temperature > 30.16728) %>% 
  distinct(Date) #89 days above?? That is 12.7 weeks

temp_data %>% 
  filter(reef == "Emerald") %>% 
  filter(Temperature > 31) 
```

## graph focusing on summertime
```{r}
str_all %>% 
  full_join(ctd_emerald_temp_sept9) %>% 
  filter(as.Date(`Datetime (EST)`) <= "2023-10-01" & as.Date(`Datetime (EST)`) >= "2023-05-31") %>% 
ggplot(., aes(x=`Datetime (EST)`, y=Temperature, color = reef)) +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "Temperature (ºC)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) +
  geom_hline(yintercept = 31, linetype = "dashed", color = "black") +
  annotate("text", x = as.POSIXct("2023-10-01"), y = 31, label = "31ºC", vjust = -1) +
  annotate("text", x = as.POSIXct("2023-09-30"), y = 33, label= "Days Above 31ºC:\nEmerald=44\nMacNorth=85", colour = "black", hjust=1) 

# Display the plot

temp_data %>% 
  select(!Date) %>% 
  filter(as.Date(`Datetime (EST)`) >= "2023-06-01" & as.Date(`Datetime (EST)`) <= "2023-10-01") %>% 
  separate(`Datetime (EST)`, into = c("Date", "Time"), sep = " ") %>% 
  mutate(Date = ymd(Date)) %>% 
  group_by(Date, reef) %>% 
  summarise(max_temp = max(Temperature)) %>% 
  filter(max_temp > 31) %>% 
  group_by(reef) %>% 
  summarise(count = n())
```

## daily variability experienced
```{r}
temp_data %>% 
  select(!Date) %>% 
  separate(`Datetime (EST)`, into = c("Date", "Time"), sep = " ") %>% 
  mutate(Date = ymd(Date)) %>% 
  group_by(Date, reef) %>% 
  summarise(daily_min = min(Temperature), daily_max=max(Temperature)) %>% 
  mutate(daily_range=daily_max - daily_min) %>% 
  ungroup() %>% 
  ggplot(., aes(x=Date, y=daily_range, color = reef)) +
  geom_point() +
  geom_line() +
  labs(x="Day", y = "Daily range (ºC)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) + 
  theme_classic()

temp_data %>% 
  select(!Date) %>% 
  separate(`Datetime (EST)`, into = c("Date", "Time"), sep = " ") %>% 
  mutate(Date = ymd(Date)) %>% 
  group_by(Date, reef) %>% 
  summarise(daily_min = min(Temperature), daily_max=max(Temperature)) %>% 
  mutate(daily_range=daily_max - daily_min) %>% 
  ungroup() %>% 
  ggplot(., aes(x=Date, y=daily_range, color = reef)) +
  geom_point() +
  geom_smooth() +
  labs(x="Day", y = "Daily range (ºC)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9")) + 
  theme_classic()
```


# PAR

# Salinity

# Tilt

load packages
```{r}
library(RColorBrewer)
library(scales)
library(lubridate)
library(openair)
```

```{r}

```


# old data analysis 

## EcoPAR

### ecopar1959 raw txt data file
```{r}
#look at raw data file 
ecopar1959_032823 <- read_tsv("environmental_data/nov22-mar23_rawdata/ecopar-sn5954-tag1959.raw.txt")

ecopar1959_032823 %>% 
  separate(col = `53651 records to read`, into = c("Date", "Time", "Value"), sep = "\t") %>% 
  mutate(Date = mdy(Date), Time = hms(Time), Value = as.numeric(Value)) -> ecopar1959_032823

length(unique(ecopar1959_032823$Date)) #115 days

ecopar1959_032823 %>% 
  filter(Date == "2022-12-31") #10 measurements were taken instantaneously at every 30 min interval for each day.

ecopar1959_032823$datetime <- paste(ecopar1959_032823$Date, ecopar1959_032823$Time)
ecopar1959_032823$datetime <- ymd_hms(ecopar1959_032823$datetime)
  
ecopar1959_032823 %>% #53,652 values
  mutate(interval = floor_date(datetime, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(Value)) %>% #5,147 values
  ggplot(., aes(x=interval, y=mean_value)) + geom_point()


# look at Mike's data tidied excel file (it looks like he made a formula to convert PAR(raw) to PAR(calc). Formula is [Digital Im] * Power(10, PARraw - [Digital a0/Digitial a1])). Unsure what the digitial things are but it changes the PAR value by approximately 10^3 (6000 -> 6)

#I found the equations from the SeaBird Scientific ECO Photosynthetically Active Radiation (PAR) sensor documentation: https://www.seabird.com/eco-photosynthetically-active-radiation-par-sensor/product-downloads?id=60762467733

#Underwater PAR
#Seasoft calculates light from an underwater Satlantic logarithmic analog PAR sensor as:
# PAR = multiplier * Im * 10 (V –a0) / a1 [user-selected units]
# where
# V = voltage output
# a0, a1, and lm are from the PAR calibration sheet that was provided by Satlantic
# multiplier = 1.0 for units of μmol photons/ m2 sec (see Notes)

#so the unit is micromol. convert to mol/day to be able to compare with Ian's paper
```



### ecopar1959 (Emerald) Dec-Mar Excel tidy data
```{r}
ecopar1959_032823_excel <-read_xlsx("Dec2022-Mar2023_rough-ecopar-sn5954-tag1959-emerald.xlsx", sheet = "sheet1")

ecopar1959_032823_excel$`Datetime (UTC)` <- as.POSIXct(ecopar1959_032823_excel$`Datetime (UTC)`, tz = "UTC")
ecopar1959_032823_excel$`Datetime (EST)` <- with_tz(ecopar1959_032823_excel$`Datetime (UTC)`, "America/New_York")

ecopar1959_032823_excel %>% #53,621 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  ggplot(., aes(x=interval, y=mean_value)) + geom_point() #values look a lot different and cleaner. I'll use these for plotting

# in Enochs et al. 2023, they "summed [PAR readings] between the hours of 10am and 3pm, to calculate a daily dose."

ecopar1959_032823_excel %>% #53,621 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  group_by(as.Date(`Datetime (EST)`)) %>% 
  mutate(hour = hour(`Datetime (EST)`), Date =as.Date(`Datetime (EST)`) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(`PAR (calc)`, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(`PAR (calc)`)) %>% 
  ggplot(., aes(x= Date , y=daily_dose)) +
  geom_point() +
  geom_line() +
  theme_classic() # these values look way higher than what Ian published as the daily dose, the range of his graph is 0-30 mol/day. I wonder if he took the mean value for the 10 instantaneous values taken at each 30 min? because adding those all together isn't an accurate representation.

ecopar1959_032823_excel %>% #53,621 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  ggplot(., aes(x= Date , y=daily_dose)) +
  geom_point() +
  geom_line() +
  theme_classic() #well that just reduced everything by an order of 10 but otherwise still way too high (~6000). That is because this is micromol. Divide by 1000 to get mol/day


ecopar1959_032823_excel %>% #53,621 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  ggplot(., aes(x= Date , y=PAR_mol_day)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "PAR (mol/day)")
```

### ecopar4298 (Emerald) Apr-Jul Excel tidy data
```{r}
ecopar4298_071123_excel <- read_xlsx("apr23-jul23_rawdata/ecopar-sn5954-tag4298_Emerald.xlsx", sheet = "ecopar-sn5954-tag4298.raw")

ecopar4298_071123_excel$`Datetime (UTC)` <- as.POSIXct(ecopar4298_071123_excel$`Datetime (UTC)`, tz = "UTC")
ecopar4298_071123_excel$`Datetime (EST)` <- with_tz(ecopar4298_071123_excel$`Datetime (UTC)`, "America/New_York")

ecopar4298_071123_excel %>% #47,040 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  ggplot(., aes(x= Date , y=PAR_mol_day)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "PAR (mol/day)")
```


### ecopar1960 (MacN) Dec-Mar Excel tidy data
```{r}
ecopar1960_032823_excel <-read_excel("Dec2022-Mar2023_rough-ecopar-sn5953-tag1960-macnorth.xlsx", sheet = "ecopar-sn5953-tag1960.raw")

ecopar1960_032823_excel$`Datetime (UTC)` <- as.POSIXct(ecopar1960_032823_excel$`Datetime (UTC)`, tz = "UTC")
ecopar1960_032823_excel$`Datetime (EST)` <- with_tz(ecopar1960_032823_excel$`Datetime (UTC)`, "America/New_York")

ecopar1960_032823_excel %>% #47,040 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  ggplot(., aes(x= Date , y=PAR_mol_day)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "PAR (mol/day)")
```

### ecopar4297 (MacN) Apr-Jul Excel tidy data
```{r}
ecopar4297_071123_excel <- read_excel("apr23-jul23_rawdata/ecopar-sn5953-tag4297_MacNorth.xlsx", sheet = "ecopar-sn5953-tag4297.raw")

ecopar4297_071123_excel$`Datetime (UTC)` <- as.POSIXct(ecopar4297_071123_excel$`Datetime (UTC)`, tz = "UTC")
ecopar4297_071123_excel$`Datetime (EST)` <- with_tz(ecopar4297_071123_excel$`Datetime (UTC)`, "America/New_York")

ecopar4297_071123_excel %>% #47,040 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  ggplot(., aes(x= Date , y=PAR_mol_day)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "PAR (mol/day)")
```

### combine all datasets into 1 so they can all be plotted together

```{r}
ecopar1959_032823_excel %>% #53,621 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  mutate(reef = "Emerald") -> ecopar1959_032823

ecopar4298_071123_excel %>% #47,040 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  mutate(reef = "Emerald") -> ecopar4298_071123 

ecopar1960_032823_excel %>% #47,040 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  mutate(reef = "MacNorth") -> ecopar1960_032823

ecopar4297_071123_excel %>% #47,040 values
  select(`Datetime (EST)`, `PAR (calc)`) %>% 
  mutate(interval = floor_date(`Datetime (EST)`, unit = "30 minutes")) %>% 
  group_by(interval) %>% 
  summarise(mean_value = mean(`PAR (calc)`)) %>% #5,369 values
  group_by(as.Date(interval)) %>% 
  mutate(hour = hour(interval), Date =as.Date(interval) ) %>% 
  ungroup() %>% 
  filter(hour >= 10 & hour <= 15) %>% 
  select(mean_value, hour, Date) %>% 
  group_by(Date) %>% 
  summarise(daily_dose = sum(mean_value)) %>% 
  mutate(PAR_mol_day = daily_dose/1000) %>% 
  mutate(reef = "MacNorth") -> ecopar4297_071123

full_join(ecopar1959_032823, ecopar4298_071123) %>% 
  full_join(., ecopar1960_032823) %>% 
  full_join(., ecopar4297_071123) -> all_ecopar_data_molperday

all_ecopar_data_molperday %>% 
  ggplot(., aes(x= Date , y=PAR_mol_day, color = reef)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "PAR (mol/day)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9"))
```

### Stats 

```{r}
all_ecopar_data_molperday

#group dates by season
#wet season = mid-may -- mid-october
#dry season = mid-october -- mid-may

all_ecopar_data_molperday %>% 
  mutate(season = case_when(Date >= "2023-05-15" ~ "Wet",
                            Date < "2023-05-15" ~ "Dry")) -> all_ecopar_data_molperday

ecopar_lm <- lm(PAR_mol_day ~ reef*season, data = all_ecopar_data_molperday)

summary(ecopar_lm)

ecopar_anova<- aov(ecopar_lm)
summary(ecopar_anova)

qqnorm(ecopar_anova$residuals) #assume normality
boxplot(resid(ecopar_anova)~all_ecopar_data_molperday$reef)
boxplot(resid(ecopar_anova)~all_ecopar_data_molperday$season) #homogeneity of variance

TukeyHSD(ecopar_anova)


```



## SeaFET



## CTD

### CTD6635 (Emerald) Dec-Mar Excel tidy data
```{r}
ctd6635_032823 <- read_xlsx("Dec2022-Mar2023_rough-ctd-emerald-6635.xlsx", sheet = "emerald-ctd-6635-data-edit")

#plot time values as is and see if the trend for time of day is weird (then it might be UTC not EST)
ctd6635_032823 %>% 
  select(Timestamp:Depth, Salinity) %>% 
  mutate(Timestamp = dmy_hms(Timestamp)) %>% 
  mutate(reef = "Emerald") -> ctd6635_032823
```

### CTD7398 (Emerald) Apr-Jul Excel data
```{r}
ctd7398_071123 <- read_xlsx("apr23-jul23_rawdata/ctd-sn7398-edited.xlsx", sheet = "ctd-sn7398-edited")

ctd7398_071123 %>% 
  select(Timestamp.UTC, Temp, Cond, Depth, Salinity) %>% 
  rename(Timestamp = Timestamp.UTC, Temperature = Temp, Conductivity = Cond) %>% 
  filter(as.Date(Timestamp) > "2023-03-28") %>% 
  mutate(reef = "Emerald") -> ctd7398_071123
```

### CTD7398 (MacN) Dec-Mar Excel
```{r}
ctd7398_032823 <- read_xlsx("Dec2022-Mar2023_rough-ctd-macnorth-7398.xlsx", sheet = "macnorth-ctd-7398-data-edit")

ctd7398_032823 %>% 
  select(Timestamp, Temperature, Conductivity, Depth, Salinity) %>% 
  mutate(Timestamp = dmy_hms(Timestamp)) %>% 
  mutate(reef = "MacNorth") -> ctd7398_032823
```

### CTD6635 (MacN) Apr-Jul 
```{r}
#data not tidied yet

#try to read the hex file
library(hexView)

hexViewFile("~/Downloads/SBE16plus_01606635_2023_07_13.hex")

#that didn't work it's not readable. needs to be run through the CTD software to extract readable numbers
```


### combine files and build temperature graph
```{r}
full_join(ctd6635_032823, ctd7398_071123) %>% 
  full_join(., ctd7398_032823) %>% 
  filter(as.Date(Timestamp) < "2023-04-01") %>% 
  ggplot(., aes(x=Timestamp, y=Temperature, color = reef)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "Temperature (ºC)") +
  scale_color_manual(values = c("#edb9c3", "#5370c9"))
```
### combine files and build salinity graph
```{r}
full_join(ctd6635_032823, ctd7398_071123) %>% 
  full_join(., ctd7398_032823) %>% 
  filter(as.Date(Timestamp) < "2023-04-01") %>% 
  filter(Salinity > 31) %>% 
  ggplot(., aes(x=Timestamp, y=Salinity, color = reef)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Day", y = "Salinity") +
  scale_color_manual(values = c("#edb9c3", "#5370c9"))
```

